
# Syntax { #Syntax }

--------------------------------------------------------------------------------

## Object names

> "There are only two hard things in Computer Science: cache invalidation and 
> naming things." 
>
> --- Phil Karlton

Variable and function names should use only lowercase letters, numbers, and underscores (`_`).    
Use `_` to separate words within a name (*snake_case style*). Using a seperator like the underscore increases readability in contrast to *lowerCamelCase* or *UpperCamelCase*.

Base R uses dots in function names (`contrib.url()`) and class names 
(`data.frame`), but it's better to reserve dots exclusively for the S3 object system. In S3, methods are given the name `function.class`; if you also use `.` in function and class names you end up with confusing methods like `as.data.frame.data.frame()`.

Generally, variable names should be nouns and function names should be verbs. 
Strive for names that are concise and meaningful (this is not easy!) and that are not too long. 

<div class = "good">
```{r, eval = FALSE}
# Good
day_one
day_1
```
</div>

<div class = "bad">
```{r, eval = FALSE}
# Bad
first_day_of_the_month
DayOne
dayone
day.one
djm1
```
</div>

* Check object names with `object_name_linter(style = "snake_case")` and `object_length_linter(length = 20L)`. 

Where possible, avoid re-using names of common functions and variables. This 
will cause confusion for the readers of your code.

<div class = "bad">
```{r, eval = FALSE}
# Bad
T <- FALSE
c <- 10
mean <- function(x) sum(x)
```
</div>

--------------------------------------------------------------------------------

## Spacing

Put a space before and after all infix operators like `=`, `+`, `-`, `<-`, `~` etc. Do without space around `^`, `/` and `*` if you feel it would increase the readability, following the principle of multiplication and division first, then addition and subtraction. Also cf. [operator precedence in R](http://stat.ethz.ch/R-manual/R-patched/library/base/html/Syntax.html).

Always put a space after a comma, and never before (just like in regular English).

<div class = "good">
```{r, eval = FALSE}
# Good
average <- mean(feet / 12 + inches, na.rm = TRUE)
average <- mean(feet/12 + inches, na.rm = TRUE)
x ^ y
sqrt(x^2 + y^2)
y ~ x
```
</div>

<div class = "bad">
```{r, eval = FALSE}
# Bad
average<-mean(feet/12+inches,na.rm=TRUE)
y~x
```
</div>

* For the comma rule there is the `commas_linter`. The `infix_spaces_linter` strictly requires spaces around all infixes (also * /), thus we don't use it.

There's another small exception to the spacing rule: don't use spaces around `:`, `::`, and `:::`.

<div class = "good">
```{r, eval = FALSE}
# Good
x <- 1:10
base::get()
```
</div>

<div class = "bad">
```{r, eval = FALSE}
# Bad
x <- 1 : 10
base :: get()
```
</div>

Extra spacing (i.e., more than one space in a row) is ok if it improves 
alignment of equal signs or assignments (`<-`).

<div class = "good">
```{r, eval = FALSE}
# Good
list(
  total = a + b + c,
  mean = (a + b + c) / n
)

list(
  total =  a + b + c,
  mean  = (a + b + c) / n
)
```
</div>

Place a space before `(`, but not when it's part of a function call. 

<div class = "good">
```{r, eval = FALSE}
# Good
if (debug) { show(x) }
plot(x, y)
```
</div>

<div class = "bad">
```{r, eval = FALSE}
# Bad
if(debug)show(x)
plot (x, y)
```
</div>

* Check these spacing rules with `spaces_left_parentheses_linter` and `function_left_parentheses_linter`.

Do not place spaces around code in parentheses or square brackets (unless 
there's a comma, in which case see above). For curly braces see section \@ref(Indenting) below.

<div class = "good">
```{r, eval = FALSE}
# Good
if (debug) { 
  do(x) 
}

diamonds[5, ]
```
</div>

<div class = "bad">
```{r, eval = FALSE}
# Bad
if ( debug ) do(x)  # No spaces around debug
x[1,]   # Needs a space after the comma
x[1 ,]  # Space goes after comma not before
```
</div>

* This can be checked with the `spaces_inside_linter`. 

--------------------------------------------------------------------------------

## Indenting {#Indenting}

Curly braces, `{}`, define the most important hierarchy of R code. To make 
this hierarchy easy to see, always indent the code inside `{}` by **two** 
spaces. A symmetrical arrangement helps with finding related braces.           
A `{` should never go on its own line and should always be followed by a new 
line.      
A `}` should always go on its own line, unless it's followed by `else` 
or `)`. 

* Check your code for this behaviour with `open_curly_linter` and `closed_curly_linter`. 
<div class = "good">
```{r, eval = FALSE}
# Good
if (y < 0 && debug) {
  message("y is negative")
}

if (x > 0) {
  message("x is positive")
} else {
  message("x is negative or zero")
}

testthat::test_that("call1 should return an ordered factor", {
  testthat::expect_s3_class(call1(x, y), c("factor", "ordered"))
})
```
</div>

<div class = "bad">
```{r, eval = FALSE}
# Bad
if (y < 0 && debug)
message("Y is negative")

if (y == 0)
{
    if (x > 0) {
      log(x)
    }else{
  message("x is negative or zero")
    }
} else { y^ x }
```
</div>

Exception: For very short `if (x) { y }` statements that fit on one line `{}` 
can go in the same line and the code inside should be surrounded by single spaces.

<div class = "good">
```{r, eval = FALSE}
# Good
x <- if (y < 20) { "Too low" } else { "Too high" }

if (y != 0) { 
  y ^ x 
} else { 
  if (x > 0) { log(x) } else { message("x is negative or zero") }
}

if (y != 0) { y ^ x } else { 
  if (x > 0) { log(x) } else { message("x is negative or zero") }
}
```
</div>

* Add argument `allow_single_line = TRUE` to the above linters to allow for the short statement exception.

Because the closing brackets belong to the same level of hierarchy as the opening brackets they should __not__ be indented. (Note that in RStudio auto-indention does not behave consistently. As a guide one can activate the _Show indent guides_ checkbox via _Global Options > Code > Display_).       
The above holds both for curly and multi-line round brackets (e.g. in function calls or definitions as in \@ref(function-head-body)).

Always use spaces for indentation, never use tabs. The difficulty with tabs is that they are not defined distinctly in terms of 
columns or spaces (a space is always one column), i.e. results would differ 
depending on the environment. 

* When using the RStudio IDE check the _Insert spaces for tab_ checkbox and specify to _Tab with 2_ in the _Global Options > Code > Editing_ tab. `no_tab_linter` also searches for tabs. 

--------------------------------------------------------------------------------

## Long lines

Strive to limit your code to __80__ characters per line. This fits comfortably 
on a printed page with a reasonably sized font. If you find yourself running out of room, this is a good indication that you should encapsulate some of the work in a separate function.

* Lint long lines with `line_length_linter(length = 80)`. 

If a function call is too long to fit on a single line, use one line each for 
the function name, each argument, and the closing `)`. This makes the code 
easier to read and to change later. 

<div class = "good">
```{r, eval = FALSE}
# Good
do_something_very_complicated(
  something = "that",
  requires = many,
  arguments = "some of which may be long"
)
```
</div>

<div class = "bad">
```{r, eval = FALSE}
# Bad
do_something_very_complicated("that", requires, many, arguments,
                              "some of which may be long"
                              )
```
</div>

You may also place several arguments on the same line if they are closely 
related to each other, e.g., strings in calls to `paste()` or `stop()`. When 
building strings, where possible match one line of code to one line of output. 

<div class = "good">
```{r, eval = FALSE}
# Good
paste0(
  "Requirement: ", requires, "\n",
  "Result: ", result, "\n"
)
```
</div>

<div class = "bad">
```{r, eval = FALSE}
# Bad
paste0(
  "Requirement: ", requires,
  "\n", "Result: ",
  result, "\n")
```
</div>

Whenever you have to write long strings that won't fit into one line, consider splitting them into smaller parts and joining those parts using `paste()`. This is not entirely elegant but conformes with the 80 char rule and ensures readability. For file paths use `file.path()`.

--------------------------------------------------------------------------------

## Assignment

Use `<-`, not `=`, for assignment.

<div class = "good">
```{r, eval = FALSE}
# Good
x <- 5
```
</div>

<div class = "bad">
```{r, eval = FALSE}
# Bad
x = 5
5 -> x 
```
</div>

* Lint `=` assignments with `assignment_linter`. 

--------------------------------------------------------------------------------

## Semicolons

Don't put `;` at the end of a line, and don't use `;` to put multiple commands 
on one line.

<div class = "bad">
```{r, eval = FALSE}
# Bad
summary(data);

x <- 5; y <- 2 * x
```
</div>

<div class = "good">
```{r, eval = FALSE}
# Good
summary(data)

x <- 5
y <- 2 * x
```
</div>

* Lint semicolons with `semicolon_terminator_linter(semicolon = c("compound", "trailing"))`. 

--------------------------------------------------------------------------------

## Quotes

Use `"`, not `'`, for quoting text. The only exception is when the text already 
contains double quotes and no single quotes. (This is really an arbitrary convention. Anyhow, please be consistent and stick to it.)

<div class = "good">
```{r, results = "hold"}
# Good
"Text"
'Text with "quotes"'
'<a href="http://style.tidyverse.org">A link</a>'

```
</div>

```{r, results = "hold"}
cat('Text with "quotes"', "\n")
cat('<a href="http://style.tidyverse.org">A link</a>')
```

<div class = "bad">
```{r, results = "hold"}
# Bad
'Text'
'Text with "double" and \'single\' quotes'
```
</div>

```{r}
cat('Text with "double" and \'single\' quotes')
```

* Lint single quotes with `single_quotes_linter`. 

--------------------------------------------------------------------------------

<div> <!-- somehow an extra closing div got inserted to the html output which resulted in missing navigation symbols (back and forth arrows)... adding the opening div statement here is a hack to solve this problem -->

