
# Functions

--------------------------------------------------------------------------------

## Naming

Use verbs for function names, where possible.

<div class = "good">
```{r eval = FALSE}
# Good
add_row()
permute()
```
</div>

<div class = "bad">
```{r eval = FALSE}
# Bad
row_adder()
permutation()
```
</div>

--------------------------------------------------------------------------------

## Head and body {#function-head-body}

If a function's arguments definition runs over multiple lines, start the definition in a second line indenting by two spaces. Every argument should go to its own line as well as the closing bracket followed by the opening curly brace `) {` and a new line for the function body.
An empty line after the opening and before the closing bracket of the body 
further increase readability.

<div class = "good">
```{r, eval = FALSE}
# Good
long_function_name <- function(
  a = "a long argument",
  b = "another argument",
  c = "another long argument"
) {
  
  # As usual code is indented by two spaces.
  
}
```
</div>

<div class = "bad">
```{r, eval = FALSE}
# Bad
long_function_name <- function(a = "a long argument",
  b = "another argument",
  c = "another long argument") {
  # Here the first argument might be overlooked 
  # and it is a bit harder to spot where the head of the function ends 
  # and the body of the function begins
}
```
</div>

--------------------------------------------------------------------------------

## Argument names

A function's arguments typically fall into two broad categories: one supplies 
the __data__ to compute on; the other controls the __details__ of computation. When 
you call a function, you typically omit the names of data arguments, because 
they are used so commonly. Specify the non-data arguments explicitly and with 
the full name, avoid partial matching.

<div class = "good">
```{r, eval = FALSE}
# Good
mean(1:10, na.rm = TRUE)
mean(1:10, trim = 0.1)
```
</div>
<div class = "bad">
```{r, eval = FALSE}
# Bad
mean(x = 1:10, , FALSE)
mean(, TRUE, x = c(1:10, NA))
mean(1:10, tr = 0.1)
```
</div>

Aside:    
Avoid abbreviations of logicals, alway use `TRUE` and `FALSE`, never use `T` 
and `F`.

<div class = "good">
```{r, eval = FALSE}
# Good
mean(1:10, na.rm = TRUE)
my_vector[my_index] <- FALSE
```
</div>
<div class = "bad">
```{r, eval = FALSE}
# Bad
mean(x = 1:10, na.rm = T)
my_vector[my_index] <- F
```
</div>

* There is a linter for spotting `T` and `F` symbols: `T_and_F_symbol_linter`. 

--------------------------------------------------------------------------------

## `return()`

Only use `return()` for early returns. Otherwise rely on R to return the result 
of the last evaluated expression.

<div class = "good">
```{r, eval = FALSE}
# Good
find_abs <- function(x, y){
  if (x > 0) { return(x) }
  -1 * x
}

add_two <- function(x, y) {
  x + y
}
```
</div>

<div class = "bad">
```{r, eval = FALSE}
# Bad
add_two <- function(x, y) {
  return(x + y)
}
```
</div>

If your function is called primarily for its side-effects (like printing, 
plotting, or saving to disk), it should return the first argument invisibly. 
(This makes it possible to use the function as part of a pipe.)     
`print` methods should usually do this, like this example from `httr`:

<div class = "good">
```{r, eval = FALSE}
# Good
print.url <- function(x, ...) {
  cat("Url: ", build_url(x), "\n", sep = "")
  invisible(x)
}
```
</div>

--------------------------------------------------------------------------------

## Comments

In code, use comments to explain the *Why* rather than the *What* or *How*. 
Each line of a comment should begin with the comment symbol and a single space: `# comment`. Your function code (the *What* and *How*) should be self-explanatory.

Comment the *What* as a headline of a codeblock optionally surrounded by lines 
of `-` or `=` to structure your script. Within functions this should rarely be necessary, as functions should not become too long. 

At [daqana](https://www.daqana.com) we agreed upon writing comments in English, and more specifically to use American English (AE), i.e. write _color_ not _colour_, use _optimize_ not _optimise_ etc. 

--------------------------------------------------------------------------------

## Design

There are two main design principles to bear in mind:

*   A function should do one thing well. 
 
    A function should be called either because it has side-effects or 
    because it returns a value; not both. Strive to keep blocks within 
    a function on one screen. 20-30 lines per function are common. For 
    functions that are significantly longer, consider splitting into smaller 
    functions.

*   A function should be easily understandable in isolation.

    Avoid global options. If your function has a transient side-effect
    (i.e. you need to create a temporary file or set an option), clean up
    after yourself with `on.exit()`.

--------------------------------------------------------------------------------
