# Pipes

--------------------------------------------------------------------------------

<!--```{block, type = "dqbox"}-->
### `r color_format("Please note", color = dq_gray)` {-}
  
The [`magrittr`](http://magrittr.tidyverse.org/) package offers the use of the pipe (`%>%`) in R. The introduction below describes situations when you might want to use the pipe.  

To begin with, here are some notes on its downsides which should be kept in 
mind. 

The `magrittr` package must be loaded and attached to the search path via a call to `library()`. The pipe, thus, must not be used in package code. I also recommend not to use it when you write functions in general as you might decide later to include them into a package.
In my opinion, the pipe should only be used in exploratory use of R, not in scripting. Here is why:  

Intermediate objects matter. You might not need them further in your analysis (that's what intermediate means), but they are crucial for spotting mistakes and for error diagnostics.
Always inspect your intermediate objects, to see if they actually look like what you expect and want them to. This is easy with named intermediate objects, but less feasible with nested function calls or pipes.

The syntax of the pipe (nice when reading out loud) conflicts with usual R syntax and requires constantly switching the mindset when used together with regular functional R syntax. Although this might just be a matter of habit, it introduces some inconsistency and thus makes your code harder to read and more error-prone.

The pipe is slightly slower than using regular function calls -- see benchmarks in our internal dqutils package.

If you decide to use the pipe anyway, follow the recommendations below.

--------------------------------------------------------------------------------

## Introduction

Use `%>%` when you find yourself composing three or more functions together 
into a nested call, or creating intermediate objects that you don't care about.
Put each verb on its own line.  This makes it simpler to rearrange 
them later, and makes it harder to overlook a step. It is ok to keep a one-step 
pipe in one line. (But then, why use the pipe?) 

* Check that each step in a pipeline is on a new line, or the entire pipe fits on one line, with `pipe_continuation_linter`. 

<div style = text-align:right>For the example cf. [Little Bunny Foo Foo](https://en.wikipedia.org/wiki/Little_Bunny_Foo_Foo)</div>

<div class = "good">
```{r eval = FALSE}
# Good
bopped <- 
  foo_foo %>%
  hop(through = forest) %>%
  scoop(up = field_mouse) %>%
  bop(on = head)

foo_foo %>% fall("asleep")     # ok
fall(foo_foo, what = "asleep") # better

hopped  <- hop(foo_foo, through = forest)
scooped <- scoop(hopped, up = field_mice)
bopped  <- bop(scooped, on = head)
# intermediate objects which are no longer needed can be cleaned up
rm(hopped, scooped)
```
</div>

<div class = "bad">
```{r eval = FALSE}
# Bad
foo_foo <- hop(foo_foo, through = forest)
foo_foo <- scoop(foo_foo, up = field_mice)
foo_foo <- bop(foo_foo, on = head)

iris %>% group_by(up) %>% summarize_all(mean) %>%
  ungroup %>% gather(field_mice, foo_foo, -on) %>%
  arrange(head)

bop( # nesting 3 or more functions gets confusing
  scoop( # nesting 2 functions is ok
    hop(foo_foo, through = forest),
    up = field_mice
  ),
  on = head
)
```
</div>

Avoid using the pipe when:

* You need to manipulate more than one object at a time. Reserve pipes for a 
  sequence of steps applied to one primary object.
 
* There are meaningful intermediate objects that could be given
  informative names.

Sometimes it seems useful to include a short pipe as an argument to a function 
in a longer pipe (which doesn't require a lookup elsewhere). Avoid this, it decreases the readability of your code. It's better to move the code outside the pipe and give it an evocative name.

<div class = "bad">
```{r eval = FALSE}
# Bad
x %>%
  select(a, b, w) %>%
  left_join(y %>% select(a, b, v), by = c("a", "b"))

x %>%
  select(a, b, w) %>%
  left_join(
    y %>%
      filter(!u) %>% 
      gather(a, v, -b) %>% 
      select(a, b, v), 
    by = c("a", "b")
    )
```
</div>

<div class = "good">
```{r eval = FALSE}
# Good
x_join <- select(x, a, b, w)
y_join <-
  y %>%
    filter(!u) %>%
    gather(a, v, -b) %>%
    select(a, b, v)
left_join(x_join, y_join, by = c("a", "b"))
```
</div>

--------------------------------------------------------------------------------

## Spacing and indenting

`%>%` should always have a space before it and a new line after it. After the 
first step, each line should be indented by two spaces.

<div class = "good">
```{r, eval = FALSE}
# Good
iris %>%
  group_by(Species) %>%
  summarize_if(is.numeric, mean) %>%
  ungroup() %>%
  gather(measure, value, -Species) %>%
  arrange(value)
```
</div>

<div class = "bad">
```{r eval = FALSE}
# Bad
iris %>% group_by(Species) %>% summarize_all(mean) %>%
ungroup %>% gather(measure, value, -Species) %>%
arrange(value)
```
</div>

--------------------------------------------------------------------------------

## No arguments

magrittr allows you to omit `()` on functions that don't have arguments. Avoid 
this.

<div class = "good">
```{r, eval = FALSE}
# Good
x %>% 
  unique() %>%
  sort()
```
</div>

<div class = "bad">
```{r eval = FALSE}
# Bad
x %>% 
  unique %>%
  sort
```
</div>

--------------------------------------------------------------------------------

## Long lines

If the arguments to a function don't all fit on one line, put each argument on 
its own line and indent:

<div class = "good">
```{r, eval = FALSE}
# Good
iris %>%
  group_by(Species) %>%
  summarise(
    Sepal.Length = mean(Sepal.Length),
    Sepal.Width = mean(Sepal.Width),
    Species = n_distinct(Species)
  )
```
</div>

--------------------------------------------------------------------------------

## Assignment

Do not use `->` to create an object at the end of the pipe. Although it 
perfectly makes sense when using the pipe (reading out loud), it conflicts with usual R syntax.

Use a separate line for the target of the assignment followed by `<-`, it makes reading the code easier. Not only is this consistent with common R syntax, also the name acts as a heading, which reminds you of the purpose of the pipe.


<div class = "good">
```{r, eval = FALSE}
# Good
iris_long <-
  iris %>%
  gather(measure, value, -Species) %>%
  arrange(-value)
```
</div>

<div class = "bad">
```{r eval = FALSE}
# Bad
iris_long <- iris %>%
  gather(measure, value, -Species) %>%
  arrange(-value)

iris %>%
  gather(measure, value, -Species) %>%
  arrange(-value) ->
  iris_long
```
</div>

--------------------------------------------------------------------------------
