# (PART) Packages {-}

# Help Pages {#help-pages}

--------------------------------------------------------------------------------

## Introduction

[Documentation of code](http://r-pkgs.had.co.nz/man.html) is essential, 
even if the only person using your code is future-you. 
Use [roxygen2](https://github.com/klutometis/roxygen) with enabled [markdown](https://github.com/klutometis/roxygen/blob/master/vignettes/markdown.Rmd)
support to keep your documentation close to the code.     
`` #` `` starts a roxygen comment, i.e. distinction from regular comments is through the backtick after the hashtag. These comments come before the object or function you are adding the documentation onto.

All help pages of *functions* should specify the following tags:

 - `@title`,
 - `@description`, 
 - `@param` (one for each function parameter/argument),
 - `@return`, 
 - `@export` (if the function is supposed to be visible outside of the package), 
 - `@author` (should be of form `firstname.lastname` at [daqana](https://www.daqana.com)) and 
 - `@examples`.
 
If applicable, also use `@family`, `@section`, `@keywords internal`.      
Nice to have: `@aliases`, `@seealso`.

```{r, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

--------------------------------------------------------------------------------

## Title and Description

For the title, describe concisely what the function does in the very 
first line of your function documentation. Titles should use sentence case 
but _not_ end with a `.` (full stop).

There is no need to use the explicit `@title` or `@description` tags, except 
in the case of the description if it is multiple paragraphs or includes 
more complex formatting like a bulleted list. I recommend to be explicit and, 
thus, to always use the tags.

<div class = "bad">
```{r}
# Bad

#' @title
#' combine values
#'

#' @title
#' Combine values into a vector or list.
#'

#' Combine values into a vector or list
#' 
#' This is a generic function which combines its arguments.
#'
```
</div>

<div class = "good">
```{r}
# Good

#' @title
#' Combine values into a vector or list
#' 
#' @description
#' This is a generic function which combines its arguments.
#'

#' @title
#' Apply a function to each element of a vector
#'
#' @description
#' The map function transform the input, returning a vector the same length
#' as the input. `map()` returns a list or a data frame; `map_lgl()`,
#' `map_int()`, `map_dbl()` and `map_chr()` return vectors of the
#' corresponding type (or die trying); `map_dfr()` and `map_dfc()` return
#' data frames created by row-binding and column-binding respectively.
#' They require dplyr to be installed.

```
</div>

--------------------------------------------------------------------------------

## Indention

Always indent with one space after `#'`. If any description corresponding to a 
`roxygen` tag spans over multiple lines, add another two spaces of extra 
indention.

<div class = "good">
```{r}
# Good

#' @param key The bare (unquoted) name of the column whose values will be used 
#'   as column headings. 
```
</div>

Alternatively, tags that span over multiple lines (like `@description`,
`@examples` and `@section`) can have the corresponding tag on its own line and 
then subsequent lines don't need extra indention.

<div class = "good">
```{r}
# Good

#' @examples
#' 1 + 1
#' sin(pi)
```
</div>

--------------------------------------------------------------------------------

## Documenting Parameters

For most tags, like `@param`, `@seealso` and `@return`, the text should be a 
sentence, starting with a capital letter and ending with a full stop.

<div class = "good">
```{r}
# Good

#' @param key The bare (unquoted) name of the column whose values will be used 
#'   as column headings. 
```
</div>

If some functions share parameters, you can use `@inheritParams` to avoid 
duplication of content in multiple places.

<div class = "tip">
```{r}
# Tip

#' @inheritParams argument function_to_inherit_from
```
</div>

--------------------------------------------------------------------------------

## Capitalization and Full Stops

For all bullets, enumerations, argument descriptions and the like, use sentence
case and put a period at the end of each text element, even if it is only a few 
words. However, avoid capitalization of function names or packages since R is 
case sensitive. Use a colon before enumerations or bulleted lists.

<div class = "good">
```{r}
# Good

#' @details In the following, we present the bullets of the list:
#' * Four cats are few animals.
#' * forcats is a package.
```
</div>

--------------------------------------------------------------------------------

## Reference

Cross-referencing is encouraged, both within R's help file system as well as to 
external resources. Include parentheses after function names when referencing 
function calls.

When referring to other sections in the documentation, use single quotes and 
title case.

<div class = "good">
```{r}
# Good

#' @seealso [fct_lump()] 

#' @param p1 Another parameter. See 'Details'.
```
</div>

<div class = "bad">
```{r}
# Bad

#' @seealso [fct_lump] 

#' @param p1 Another parameter. See "details".
```
</div>

If you have a family of related functions, you can use the `@family` tag with a 
shared family name to automatically add appropriate lists and interlinks to the 
`@seealso` section. Family names are plural.   
In `dplyr`, the verbs `arrange()`, `filter()`, `mutate()`, `slice()`, 
`summarize()` form the family of single table verbs. Each documentation contains
the following line:

<div class = "tip">
```{r}
# Tip

#' @family single table verbs
```
</div>

--------------------------------------------------------------------------------

## Line Break

Leave one line blank before / after each description.

<div class = "good">
```{r}
# Good

#' @section Tidy data:
#' When applied to a data frame, row names are silently dropped. To preserve,
#' convert to an explicit variable with [tibble::rownames_to_column()].
#'
#' @section Scoped filtering:
#' The three [scoped] variants ([filter_all()], [filter_if()] and
#' [filter_at()]) make it easy to apply a filtering condition to a
#' selection of variables.
```
</div>

If you want to insert a line break within a section, also leave a line break in 
the roxygen comments.

<div class = "tip">
```{r}
# Tip

#' @param ... Data frames to combine.
#'
#'   Each argument can either be a data frame, a list that could be a data
#'   frame, or a list of data frames.
#'
#'   When row-binding, columns are matched by name, and any missing
#'   columns with be filled with NA.
#'
#'   When column-binding, rows are matched by position, so all data
#'   frames must have the same number of rows. To match by value, not
#'   position, see [join()].
```
</div>

--------------------------------------------------------------------------------

## Code Font

Text that contains valid R code should be marked as such using `` `code` `` (backticks around `code`).

* Function names, should be followed by `()`, e.g. `tibble()`.
* Function arguments, e.g. `na.rm`. 
* Values, e.g. `TRUE`, `FALSE`, `NA`, `NaN`, `...`, `NULL` 
* Literal R code, e.g. `mean(x, na.rm = TRUE)`

To refer to the R language, use `\R`.
Do not use code font for package names. If package name is ambiguous in the 
context, disambiguate with words, e.g. "the foo package".


--------------------------------------------------------------------------------

## Internal Functions

Internal functions should be documented with `#'` comments as per usual. 
Use the `@noRd` tag to prevent `.Rd` files from being generated.

<div class = "good">
```{r}
# Good

#' @title Drop last
#'
#' @description 
#' Drops the last element from a vector.
#'
#' @param x A vector object to be trimmed.
#'
#' @noRd
```
</div>

--------------------------------------------------------------------------------

## Deprecating Functions

The following example is taken from https://stackoverflow.com/a/10145627.
<!--(Also see answer https://stackoverflow.com/a/14036524 below.)-->

<div class = "good">
```{r}
# Good

## before

#' @title Alchemy function with too long name
#' @description Turn lead into gold.
#' @param lead Input material.
#' @return gold
#' @export
transmute_my_carefully_crafted_data_structure_into_gold <- function(lead) {
  # todo: figure out how to create gold
  # look like we are doing something
  Sys.sleep(10)
  return("gold")
}

## after

#' @title Alchemy function with too long name
#' @description Deprecated function. Turn lead into gold.
#' @param lead Input material.
#' @return gold
#' @export
#' @rdname package-deprecated
#' @seealso \code{\link{alchemy}}
transmute_my_carefully_crafted_data_structure_into_gold <- function(lead) {
  .Deprecated("alchemy", package = "your_package_name")
  alchemy(lead)
}

#' @title Alchemy function with appropriate name
#' @description Turn lead into gold.
#' @param lead Input material.
#' @return gold
#' @export
alchemy <- function(lead) {
  # todo: figure out how to create gold
  # look like we are doing something
  Sys.sleep(10)
  return("gold")
}
```
</div>

--------------------------------------------------------------------------------
