# Error messages

--------------------------------------------------------------------------------

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

An error message should start with a general statement of the problem and then give
a concise description of what went wrong. Consistent use of punctuation and
formatting makes errors easier to parse.

--------------------------------------------------------------------------------

## Problem statement

Every error message should start with a general statement of the problem. It
should be concise, but informative. (This is hard!)

* If the cause of the problem is clear use `must`. Clear cut causes typically involve incorrect types or lengths.

```{r}
dplyr::nth(1:10, "x")
```
<div class = "good">
```{r}
# GOOD
## Error: `n` must be a numeric vector, not a character vector.
```
</div>

```{r}
dplyr::nth(1:10, 1:2)
```
<div class = "good">
```{r}
# GOOD
## Error: `n` must have length 1, not length 2.
```
</div>

* If you cannot state what was expected, use `can't`:

```{r}
mtcars %>% pull(b)
```
<div class = "good">
```{r}
# GOOD
## Error: Can't find column `b` in `.data`.
```
</div>

```{r}
as_vector(environment())
```
<div class = "good">
```{r}
# GOOD
## Error: Can't coerce `.x` to a vector.
```
</div>

```{r}
purrr::modify_depth(list(list(x = 1)), 3, ~ . + 1)
```
<div class = "good">
```{r}
# GOOD
## Error: Can't find specified `.depth` in `.x`.
```
</div>

Use `stop(call. = FALSE)`, `rlang::abort()`, `Rf_errorcall(R_NilValue, ...)` to
avoid cluttering the error message with the name of the function that generated
it. That information is often not informative, and can easily be accessed via
`traceback()` or an IDE equivalent.

--------------------------------------------------------------------------------

## Error location

Do your best to reveal the location, name, and/or content of the troublesome
component. The goal is to make it as easy as possible for the user to find and fix
the problem.

```{r}
map_int(1:5, ~ "x")
```
<div class = "bad">
```{r}
# BAD
## Error: Each result must be a single integer
```
</div>
<div class = "good">
```{r}
# GOOD
## Error: Each result must be a single integer:
## * Result 1 is a character vector.
```
</div>

(It is often not easy to identify the exact problem; it may require passing
around extra arguments so that error messages generated at a lower-level can
know the original source. For frequently used functions, the effort is typically
worth it.)

If the source of the error is unclear, avoid pointing the user in the wrong
direction by giving an opinion about the source of the error:

```{r}
pull(mtcars, b)
```
<div class = "good">
```{r}
# GOOD
## Error: Can't find column `b` in `.data`.
```
</div>
<div class = "bad">
```{r}
# BAD: implies one argument at fault
## Error: Column `b` must exist in `.data`

## Error: `.data` must contain column `b`
```
</div>

```{r}
tibble(x = 1:2, y = 1:3, z = 1)
```
<div class = "good">
```{r}
# GOOD
## Error: Columns must have consistent lengths: 
## * Column `x` has length 2
## * Column `y` has length 3
```
</div>
<div class = "bad">
```{r}
# BAD: implies one argument at fault
## Error: Column `x` must be length 1 or 3, not 2 
```
</div>

If there are multiple issues, or an inconsistency revealed across several
arguments or items, prefer a bulleted list:

```{r}
purrr::reduce2(1:4, 1:2, `+`)
```
<div class = "good">
```{r}
# GOOD
## Error: `.x` and `.y` must have compatible lengths:
## * `.x` has length 4
## * `.y` has length 2
```
</div>
<div class = "bad">
```{r}
# BAD: harder to scan
## Error: `.x` and `.y` must have compatible lengths: `.x` has length 4 and 
## `.y` has length 2
```
</div>

--------------------------------------------------------------------------------

## Hints

If the source of the error is clear and common, you may want provide a hint as
how to fix it:

```{r}
dplyr::filter(iris, Species = "setosa")
```
<div class = "good">
```{r}
# GOOD
## Error: Filter specifications must be named.
## Did you mean `Species == "setosa"`?
```
</div>

```{r}
ggplot2::ggplot(ggplot2::aes())
```
<div class = "good">
```{r}
# GOOD
## Error: Can't plot data with class "uneval". 
## Did you accidentally provide the results of aes() to the `data` argument?
```
</div>

Hints should always end in a question mark.

Hints are particularly important if the source of the error is far away from the
root cause:

```{r}
mean[[1]]
```
<div class = "bad">
```{r}
# BAD
## Error in mean[[1]] : object of type 'closure' is not subsettable.
```
</div>
<div class = "good">
```{r}
# BETTER
## Error: Can't subset a function.

# BEST
## Error: Can't subset a function.
## Have you forgotten to define a variable named `mean`?
```
</div>

Good hints are difficult to write because, as above, you want to avoid steering
users in the wrong direction. Generally, I avoid writing a hint unless the
problem is common, and you can easily find a common pattern of incorrect usage
(e.g. by searching StackOverflow).

--------------------------------------------------------------------------------

## Grammar

* Errors should be written in sentence case, and should end in a full stop.
  Bullets should be formatted similarly; make sure to capitalise the first
  word (unless it's an argument or column name).
  
* Prefer the singular in problem statements:

```{r}
map_int(1:2, ~ "a")
```
<div class = "bad">
```{r}
# BAD
## Error: Results must be coercible to single integers: 
## * Result 1 is a character vector
```
</div>
<div class = "good">
```{r}
# GOOD
## Error: Each result must be coercible to a single integer:
## * Result 1 is a character vector.
```
</div>

* If you can detect multiple problems, list up to five. This allows the user
  to fix multiple problems in a single pass without being overwhelmed by
  many errors that may have the same source.

<div class = "good">
```{r}
# BETTER
## Error: Each result must be coercible to a single integer:
## * Result 1 is a character vector
## * Result 2 is a character vector
## * Result 3 is a character vector
## * Result 4 is a character vector
## * Result 5 is a character vector
## * ... and 5 more problems
```
</div>

* Pick a natural connector between problem statement and error location:
  this may be `, not`, `;`, or `:` depending on the context.

* Surround the names of arguments in backticks, e.g. `` `x` ``. 
  Use `column` to disambiguiate columns (statistical variables) and arguments: `` Column `x` ``.
  Avoid the term `variable`, because it is ambiguous (statistics vs. computer science).
 
* Ideally, each component of the error message should be less than 80 
  characters wide. Do not add manual line breaks to long error messages; 
  they will not look correct if the console is narrower (or much wider) than 
  expected. Instead, use bullets to break up the error into shorter logical
  components.

--------------------------------------------------------------------------------

## Examples

The process of improving error messages is illustrated below through _before_ and _after_ examples from around the tidyverse.

```{r}
dplyr::filter(mtcars, cyl)
```
<div class = "bad">
```{r}
# before:
## Argument 2 filter condition does not evaluate to a logical vector
```
</div>
<div class = "good">
```{r}
# after:
## Each argument must be a logical vector:
## * Argument 2 (`cyl`) is an integer vector.
```
</div>

```{r}
tibble::tribble("x", "y")
```
<div class = "bad">
```{r}
#before:
## Expected at least one column name; e.g. `~name`
```
</div>
<div class = "good">
```{r}
# after:
## Must supply at least one column name, e.g. `~name`.
```
</div>

```{r}
ggplot2::ggplot(data = diamonds) + 
  ggplot2::geom_line(ggplot2::aes(x = cut))
```
<div class = "bad">
```{r}
#before:
## geom_line requires the following missing aesthetics: y
```
</div>
<div class = "good">
```{r}
# after:
## `geom_line()` must have the following aesthetics: `y`.
```
</div>

```{r}
dplyr::rename(mtcars, cyl = xxx)
```
<div class = "bad">
```{r}
#before:
## `xxx` contains unknown variables
```
</div>
<div class = "good">
```{r}
# after:
## Can't find column `xxx` in `.data`.
```
</div>

```{r}
dplyr::arrange(mtcars, xxx)
```
<div class = "bad">
```{r}
#before:
## Evaluation error: object 'xxx' not found.
```
</div>
<div class = "good">
```{r}
# after:
## Can't find column `xxx` in `.data`.
```
</div>

--------------------------------------------------------------------------------
